My title

### 2024-09-23 paper/paper.qmd
draw some graph

library(tidyverse)

@@

cleaned_shelter_data <- read.csv("/Users/liz/Downloads/starter_folder-main/data/analysis_data/analysis_data.csv")

# Ensure the date column is in the correct format (Month-Year format)
cleaned_shelter_data$date <- as.Date(paste0("01-", cleaned_shelter_data$date), format="%d-%b-%y")

# Filter to include only the "All Population" group
total_population_data <- subset(cleaned_shelter_data, group == "All Population")

# Load ggplot2 for plotting
ggplot(total_population_data, aes(x = date, y = homeless_count)) +
  geom_line(color = "blue", linewidth = 1.5) +  
  labs(title = "Total Homeless Population Over Time",
       x = "Date",
       y = "Homeless Population Count") +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +  
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

@@

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

>>>>

library(tidyverse)
library(ggplot2)
library(dplyr)
library(scales) 
library(tidyr)

@@

cleaned_shelter_data <- read.csv("/Users/liz/Downloads/starter_folder-main/data/analysis_data/shelter_analysis_data.csv")

# Ensure the date column is in the correct format (Month-Year format)
cleaned_shelter_data$Date <- as.Date(paste0("01-", cleaned_shelter_data$Date), format="%d-%b-%y")

# Filter to include only the "All Population" group
total_population_data <- subset(cleaned_shelter_data, Population_Group == "All Population")

# Load ggplot2 for plotting
ggplot(total_population_data, aes(x = Date, y = Homeless_Count)) +
  geom_line(color = "blue", linewidth = 1.5) +  # Adjust line color and width
  labs(title = "Total Homeless Population Over Time",
       x = "Date",
       y = "Homeless Population Count") +
  scale_x_date(date_labels = "%Y", date_breaks = "1 year") +  # Show year on x-axis with 1-year intervals
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

@@

```{r}

# Summarize data for all years (total homeless population by age group)
total_data <- cleaned_shelter_data %>%
  summarise(Age_Under_16 = sum(Age_Under_16, na.rm = TRUE),
            Age_16_24 = sum(Age_16_24, na.rm = TRUE),
            Age_25_34 = sum(Age_25_34, na.rm = TRUE),
            Age_35_44 = sum(Age_35_44, na.rm = TRUE),
            Age_45_54 = sum(Age_45_54, na.rm = TRUE),
            Age_55_64 = sum(Age_55_64, na.rm = TRUE),
            Age_65_Over = sum(Age_65_Over, na.rm = TRUE))

# Reshape data for plotting
total_data_long <- tidyr::pivot_longer(total_data, 
                                       cols = starts_with("Age"), 
                                       names_to = "Age_Group", 
                                       values_to = "Count")

# Calculate percentages
total_count <- sum(total_data_long$Count)
total_data_long <- total_data_long %>%
  mutate(Percent = round(Count / total_count * 100))

# Create a pie chart with percentage labels
ggplot(total_data_long, aes(x = "", y = Count, fill = Age_Group)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y") +
  geom_text(aes(label = paste0(Percent, "%")), 
            position = position_stack(vjust = 0.5), size = 4) +
  labs(title = "Age Group Distribution of Total Homeless Population") +
  theme_void() +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")  # Use a color palette for clarity

```{r}

# Read the cleaned data
cleaned_shelter_data <- read.csv("/Users/liz/Downloads/starter_folder-main/data/analysis_data/shelter_analysis_data.csv")

# Ensure the date column is in the correct format
cleaned_shelter_data$date <- as.Date(paste0("01-", cleaned_shelter_data$Date), format="%d-%b-%y")
cleaned_shelter_data$year <- format(cleaned_shelter_data$date, "%Y")  # Extract year as a string

# Summarize data by year for different age groups
age_data <- cleaned_shelter_data %>%
  group_by(year) %>%
  summarise(Age_Under_16 = sum(Age_Under_16, na.rm = TRUE),
            Age_16_24 = sum(Age_16_24, na.rm = TRUE),
            Age_25_34 = sum(Age_25_34, na.rm = TRUE),
            Age_35_44 = sum(Age_35_44, na.rm = TRUE),
            Age_45_54 = sum(Age_45_54, na.rm = TRUE),
            Age_55_64 = sum(Age_55_64, na.rm = TRUE),
            Age_65_Over = sum(Age_65_Over, na.rm = TRUE))

# Reshape data for plotting
age_data_long <- pivot_longer(age_data, 
                              cols = starts_with("Age"), 
                              names_to = "Age_Group", 
                              values_to = "Count")

# Plot 100% stacked bar chart
ggplot(age_data_long, aes(x = year, y = Count, fill = Age_Group)) +
  geom_bar(stat = "identity", position = "fill") +  # Use 'fill' to normalize to 100%
  scale_y_continuous(labels = percent) +  # Display percentages on y-axis
  labs(title = "Age Group Proportion Over Time", 
       x = "Year", y = "Proportion", fill = "Age Group") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")  # Use color palette for clarity

@@

# Read the cleaned data
cleaned_shelter_data <- read.csv("/Users/liz/Downloads/starter_folder-main/data/analysis_data/shelter_analysis_data.csv")

# Ensure the date column is in the correct format
cleaned_shelter_data$date <- as.Date(paste0("01-", cleaned_shelter_data$Date), format="%d-%b-%y")
cleaned_shelter_data$year <- as.numeric(format(cleaned_shelter_data$date, "%Y"))  # Extract year as numeric

# Summarize data by year for different genders
gender_data <- cleaned_shelter_data %>%
  group_by(year) %>%
  summarise(Male = sum(Male_Count, na.rm = TRUE),
            Female = sum(Female_Count, na.rm = TRUE),
            Non_Binary = sum(Transgender_Non_Binary_Count, na.rm = TRUE))

# Calculate total population for each year and for overall gender proportions
total_population <- gender_data %>%
  summarise(Male = sum(Male), Female = sum(Female), Non_Binary = sum(Non_Binary))

# Calculate percentages for each gender over the entire period
total_population_long <- pivot_longer(total_population, 
                                      cols = c("Male", "Female", "Non_Binary"), 
                                      names_to = "Gender", 
                                      values_to = "Total_Count")

# Calculate proportions
total_population_long <- total_population_long %>%
  mutate(Percent = round(Total_Count / sum(Total_Count) * 100, 1))

# Create a string for the percentages to use in the legend
gender_percentages <- paste0(total_population_long$Gender, " (", total_population_long$Percent, "%)")

# Reshape data for plotting
gender_data_long <- pivot_longer(gender_data, 
                                 cols = c("Male", "Female", "Non_Binary"), 
                                 names_to = "Gender", 
                                 values_to = "Count")

# Ensure gender order: Non-Binary at the bottom, Female in the middle, Male at the top
gender_data_long$Gender <- factor(gender_data_long$Gender, levels = c("Non_Binary", "Female", "Male"))

# Plot area chart with percentages in the legend
ggplot(gender_data_long, aes(x = year, y = Count, fill = Gender)) +
  geom_area(position = "stack") +
  labs(title = "Homeless Population by Gender Over Time (Area Chart)", 
       x = "Year", y = "Homeless Population Count", fill = "Gender") +
  scale_fill_manual(values = c("Male" = "blue", "Female" = "pink", "Non_Binary" = "purple"),
                    labels = gender_percentages) +  # Add percentages to the legend
  theme_minimal()

Talk way more about it. 
```{r}

# Summarize data by year for Chronic and All Population
chronic_data <- cleaned_shelter_data %>%
  filter(Population_Group == 'Chronic') %>%
  group_by(year) %>%
  summarize(homeless_count = sum(Homeless_Count))

all_population_data <- cleaned_shelter_data %>%
  filter(Population_Group == 'All Population') %>%
  group_by(year) %>%
  summarize(homeless_count = sum(Homeless_Count))

# Create Non-Chronic by subtracting Chronic from All Population
non_chronic_data <- all_population_data
non_chronic_data$homeless_count <- non_chronic_data$homeless_count - chronic_data$homeless_count
non_chronic_data$group <- 'Non-Chronic'

# Add Chronic to the dataset
chronic_data$group <- 'Chronic'
combined_data <- rbind(chronic_data, non_chronic_data)  # Combine data

# Ensure 'group' is a factor and order it so that Chronic appears at the top
combined_data$group <- factor(combined_data$group, levels = c("Non-Chronic", "Chronic"))

# Reorder the data frame so that 'Non-Chronic' is plotted first (to appear at the bottom)
combined_data <- combined_data %>%
  arrange(group)

# Plot the stacked bar chart with Chronic at the top
ggplot(combined_data, aes(x = year, y = homeless_count, fill = group)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(title = "Chronic vs Non-Chronic Homeless Population by Year", 
       x = "Year", y = "Homeless Population Count", fill = "Group") +
  theme_minimal() +
  scale_fill_manual(values = c("Chronic" = "blue", "Non-Chronic" = "orange"))  # Adjust colors if needed
